---
title: "DunbarRae_Ch2_Part1.Rmd"
output: html_notebook
---
### Initial Setup for R

getting pckages in that are needed.
```{r}
install.packages("dplyr")
library(dplyr)
library(ggplot2)
```


```{r}
rm(list=ls()) #Clear R's Brain
```

To get data into R there are a few methods:

### Method 1
```{r}
#This is method 1
library(readr)
compensation_2_ <- read_csv("Data for the code/compensation (2).csv")
View(compensation_2_)
```

We know this method is not great, as when you clear R's brain the file directory
is lost. So, we want to actually write out the entire correct direcotry. Or
the R book mentions a different way. 

### Method 2:
this is much more accurate and simpler. 
```{r}
#use file.choose() in your consol to get the file directory
compensation<-read.csv("/home/rldunb19/Biostats Senior Fall/Rae-Homework/Data for the code/compensation (2).csv") #fllows the file directory and reads it into R, saves
#it under the compensation lable. 
```

### Method 3:
Now I went to the "session" column up above by plots and Build and then clicked
"Set working directory" then "choose directory" and chose my Data file. Now this should work:

```{r}
compensationM1<-read.csv("compensation (2).csv")
```
This method did not work, which is a mega bummer :(. Do you know why Erika??

### Method 4
you can also put your data in the same place where you put your script/Rmd file. 
Why don't we do this?

### Checking your Data

```{r}
#to get the names of the columns:
names(compensation)

# should return the first 6 rows
head(compensation)

# should return the last 6 rows
tail(compensation)

# number of rows and columns
dim(compensation)

#returns the structure of the data
str(compensation)
```

### Using dyplr
Two functions from dplyr are also use- ful for looking at the data you’ve just imported.
```{r}
# look at your data in a horizontal view
glimpse(compensation)

# provides a vertical view
tbl_df(compensation)
```
