---
title: "Practicing Chp. 1-3"
author: "Rae Dunbar"
date: "10/13/2022"
output: html_document
---
## Setting up R
```{r}
rm(list=ls())
library(dbplyr)
library(here)
library(ggplot2)
```


```{r}
leaves<- read.csv(here("Data for the code", "Class_leaf_measurements.csv"), stringsAsFactors = TRUE)

```

```{r}
str(leaves)
```
this tells me I forgot to do stringsasFactors as they are listed as characters, so I went back up and did that. I also see some dots in the column headings, which need to be fixed. 

## Fixing the column names

lets see what we are dealing with.
```{r}
names(leaves) #shows the column headers
```

lets now fix them using the names function. 
```{r}
names(leaves)<-c("Tree_num", "Species" , "Leaf_num" ,  "DBH_cm", "Petiole_length" , "Total_length_cm", "Middle_lobe_length_cm" , "Middle_lobe_width_cm", "Person" )
```

double checking. 
```{r}
names(leaves)
```

## Fixing Species

species is already a factor variable as I did that above, so I can use levels to see what the data variations  are.
```{r}
levels(leaves$Species)
```
We have 3 norway maples, one spelled wrong, and with a capital m and a lowercase m. Then we have 3 red maples  and 3 sugar maples. So we will have to fix all of those as well. 

### Fixing the norways
```{r}
fix<- which(leaves$Species== "Norway mapel" | leaves$Species== "Norway maple" )
leaves$Species[fix]<- "Norway Maple" # make sure to use the brackets and not the ()! 
```

### Fixing the reds

```{r}
fixred<- which(leaves$Species== "red maple" | leaves$Species== "Red maple")
leaves$Species[fixred]<- "Red Maple" #if I did fix again as my varialbe, we may run into some problems later on if we only run parts of this rmd. 
```
I used Red Maple for datum consistency. 

### Fixing the sugars

```{r}
fixsugar<- which(leaves$Species=="sugar maple" | leaves$Species== "Sugar mapel")
leaves$Species[fixsugar]<- "Sugar Maple"
```

```{r}
levels(leaves$Species)
```
this is annoying R thing. As all of the incorrect spellings are still options for that column, so they still come up as 'levels.' But if you check the data the code worked.

HERE IS THE CODE THAT FIXES THE LEVELS AFTER YOU DO THE RELEVELING:

```{r}
leaves$Species<-factor(leaves$Species)
levels(leaves$Species)
```


## Selecting just the Sugar Maples

```{r}
sugar<- leaves %>% filter(Species=="Sugar Maple")
```
This was the oddest thing, I ran it the first time and it did not work. I ran all my code again and it still did not work. I ran another homework set and then reran this one and it worked. What the heck!

I BELIEVE THE REASON FOR THIS PROBLEM IS THAT YOU HAVE A TYPO IN LINE 10 WITH `library(dplyr)' IN WHICH YOU SPELLED `dplyr` INCORRECTLY. THUS, THE CORRECT FILTER FUNCTION IS NOT LOADED. TRY FIXING IT AND RUN AGAIN.


## remove undesired columns
```{r}
sugar<- sugar %>% select(-"Person", -"Leaf_num", -"Tree_num") #select is for columns, slice is for rows, I keep mixing those up!
```

## Make a histogram??
I do not think we learned this...

WE SAW IT IN CHAPTER 2 SO I PUT IT HERE AGAIN.
```{r}
geom_histogram(sugar$petiole_length)
```
Rae's rusty attempt at a histogram. 

The book's correct methodology. 
```{r}
ggplot(sugar, aes(x=Petiole_length)) + geom_histogram()
# use the ggplot function to take sugar and then aes (wish I knew what this stood for) so the x axis is petiole length. then you tell it to make a histogram. 
```
It is not the prettiest on the block. Soon I reckon we will learn how to spruce it up a bit and get it ready for prom.

HOORAY FOR HISTOGRAM PROM!
