---
title: "Practice with dplyr"
author: "Rae Dunbar"
date: "9/26/2022"
output: html_document
---
Practice for Chapters 2 and 3 of the R programming book. 
## Setting up R
```{r}
rm(list=ls()) #clear R's brain
```

add the packages
```{r}
library(dplyr)
library(here)
library(ggplot2)
```

Getting in the data

```{r}
Cover<- read.csv(here("Data for the code", "donnerville_percent_cover.csv"), stringsAsFactors = TRUE)
#stringsAsFactors=TRUE is important for statistical analysis.
```

## Examing the Data

Lets see what this data is.

```{r}
names(Cover) #return the column names
```
I would consider renaming the column header Plot.dim to something that is more intuitive, as dim could be diameter or dimension. I am guessing it is diameter given the plot.type is a circle, however so changing is not necessary. The other column headers look good.

renaming column headings, to fix the . in there which means there was a space in the csv.
```{r}
Cover<- Cover %>% rename(Plot_type= Plot.type, Plot_diameter=Plot_dim) 
```

another way
```{r}
names(Cover)<- c("site","Date", "Plot_type", "Plot_diameter", "Slope", "Aspect", "Plot", "type", "Percent")
```


```{r}
head(Cover) #return the first 6 rows of the data
```
Now that I have a good idea as to what this data is, lets check our data to ensure there are no errors.

```{r}
summary(Cover)
```

The mean, median, 1st and 3rd quartile numbers all look reasonable, there appears to be no concerning outliers that suggest improper data recording. I also notice there are several NAs for Aspect, but not for any other column. 

## Manipulating the Data

9. Determine the average percent cover by cover type (bare, bryophytes, vascular plants) for Donnerville 1. 
I initially forgot that we only wanted Donnerville 1, oops.
```{r}
Cover %>% filter(Site=="Donnerville 1") %>% group_by(Type) %>% summarise(avgPercCover= mean(Percent)) 
# take the data frame 'Cover' and then filter so you only have the site Donnerville 1, and then group by the column Type and then make a new column called avgPerceCover which will have the mean of the percent cover per type. 
```
10.Determine the average percent vascular plant cover across all sites.

So, site does not matter, we want the average percent cover of vascular plants in all the samples.  
```{r}
Cover %>% filter(Type=="Vascular") %>% summarise(avgPercCover=mean(Percent))
# take the data frame 'Cover' and then filter out to only include the vacular type and then make a new column called avgPerceCover which will have the mean of the percent cover of vascular plants.
```
Yay :)
give Gus a treat for me.
