---
title: 'Making Graphs: Whisker, Histogram,'
author: "Rae Dunbar"
date: "10/24/2022"
output: html_document
---

## Setting up R
```{r}
rm(list=ls())
library(dplyr)
library(here)
library(ggplot2)
```

## Getting in the data
```{r}
soil<- read.csv(here("Data for the code", "soil_data.csv"))
```

## Checking out the data
```{r}
View(soil)
```

## Diving right in to make the figures

And we get this mess:
```{r}

ggplot(soil, aes(x=SOM_percent, y=Field.Name)) + theme_bw() +geom_point()
```

## A Sad Realization

So we have a big issue, the forest where each sample was taken is in a column with other data we do not need. Come on people, we are suppose to put one datum per entry! 

We want our data per forest, so Edward and I in the pub are going to put our heads together and figure it out. 

## The Power of 2
```{r}
soil2<- mutate(soil,forest=NA) #we need to create a new column called forest.I just filled it in with NA for now.
```

To back fill it with the proper forest names we could do which() though that would be tedious. I decided to just fill them in manually of sorts, with 5 lines of code. I bet there is a quicker way that I hope you show me in class. I wonder if you can do it in the mutate code somehow. I did use concatenate to try and do that, but it needed me to write each forest name 9 times. Is there a way aside from a loop? 
```{r}
#The code below fills in the data from row 1 to 9 in the forest column.
soil2[1:9,"forest"]<- "Degrasse" 
soil2[10:18,"forest"]<-"Donnerville"
soil2[19:27,"forest"]<-"Kip Tract"
soil2[28:36,"forest"]<- "Peavine"
soil2[37:45,"forest"]<- "South Hammond"

glimpse(soil2)
```

At first, the glimpse showed me that south Hammond was filled in as the first few entries, which is wrong. I made the error of doing 1-9 rather than 1:9. Now it seems to work.

## Making Figures finally!!!
```{r}
ggplot(soil2, aes(x=forest, y=SOM_percent)) +theme_bw() + geom_point() 
```
You could do colour=Field.Name to show each plot, but you would have to get rid of the key for that graph, as the key is giant! I do not know how to get rid of the key, so I left it black and white for now. 

```{r}
ggplot(soil2, aes(y=SOM_percent, colour=forest))+theme_bw() +geom_boxplot()
#this one I did a big differently, by differentiating between the forest treatment by color. 
```
The vast range in the peavine forest is very interesting, given the small range in the other forests. There is also a noticeable contract in SOM of Peavine when compared to Degrasse and South Hammond. 

Doing the color method rather than having forest on the x axis, causes the x axis to randomly have numbers which makes no sense, so I will not do that in the future.

```{r}
ggplot(soil2, aes(x=SOM_percent, colour=forest)) + theme_bw() +geom_histogram()
```
This is weird.... I think making several histograms per forest might make fore sense. 

```{r}
ggplot(soil2, aes(x=SOM_percent)) + theme_bw() +geom_histogram(bins = 10)+facet_wrap(~forest) #bins increases the width of the boxes so it is easier to see. 
```

This looks a bit better, though I still do not like histograms. 

## On to Calcium!

```{r}
ggplot(soil2, aes(x=forest, y=BaseSaturationCa)) + xlab("Forest Type") + ylab("Calcium ppm") +theme_bw() + geom_point() 
# You can change you x and y labels by adding more layers to your graph with xlab and ylab
```
```{r}
ggplot(soil2, aes(y=BaseSaturationCa, x=forest))+theme_bw()+ xlab("Forest Type") + ylab("Calcium ppm") +geom_boxplot()
```
```{r}
ggplot(soil2, aes(x=BaseSaturationCa)) + theme_bw() + xlab("Forest Type") + ylab("Calcium ppm") +geom_histogram(bins = 10)+ facet_wrap(~forest) #bins increases the width of the boxes so it is easier to see. 
```


